---
title: "UMD Fees"
date: 8-25-2025
---

Follow along with the demo to learn how to calculate percentage change using R and the Tidyverse. You should get started by loading the tidyverse in a new R block.


### Task 1: Load the tidyverse, janitor
Janitor performs important data cleaning operations

```{r}
library(tidyverse)
#install.packages('janitor')
library(janitor)
```




### Task 2: Load the UMD Fees dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/umd_fees.csv

```{r}
umd_fees <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/umd_fees.csv") |> 
   clean_names()
```

### Task 3: Calculate the percentage change in each category between Fall 2024 and Fall 2021 using `mutate()`

```{r}
umd_fees_pct <- umd_fees |> mutate(pct_change = (fall_2024- fall_2021)/fall_2021 * 100)
```


### Task 4: Calculate the mean and median fees for Fall 2024 using `mutate()`

```{r}

umd_fees |> 
    summarize(mean_2024_fees = mean(fall_2024),
            median_2024_fees = median(fall_2024)
)


```

### Task 5: Add to the previous the max and minimum values for 2024 in addition to the mean and median fees for Fall 2024 using `mutate()`

```{r}

umd_fees |> 
    summarize(mean_2024_fees = mean(fall_2024),
            median_2024_fees = median(fall_2024),
            max_2024_fees = max(fall_2024),
            min_2024_fees = min(fall_2024)
)


```

### Task 6: Describe our results

Write a few sentences about what you found in the data. What's the story here?





When you are done, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
