---
title: "Diversity Index for Maryland Cities"
date: "Nov 8, 2025"
author: "Your Name"
---

Wesley Schnell


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Overview

In this exercise, we will calculate the **USA TODAY Diversity Index** for Maryland cities with a population of at least 10,000 using 2020 Census race and ethnicity data. The Diversity Index reflects the probability that two randomly selected individuals from a city are of different races or ethnicities.

A score of 0 indicates no diversity (everyone is the same race/ethnicity), while a score closer to 100 indicates higher diversity. The **most diverse city** has the highest index, reflecting a balanced mix of several racial and ethnic groups. The **least diverse** city likely has one overwhelmingly dominant racial/ethnic group.

## Task 1: Load and Describe the Data (3 points)

Run the following code to load the data, then examine it and describe what it contains. Based on the raw numbers alone, which cities stand out as diverse to you?



```{r load-data}
md_cities <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/usat_diversity_index.csv")

# Preview structure
View(md_cities)
```

Baltimore city stands out to me as diverse because they have the highest population of all cities, which means there are going to be a bunch of different people for each race. Colombia and Germantown are in the same boat as they have high populations, therefore, leading to a more diverse city. 


## Task 2: Calculate Race Percentages

```{r race-percentages}
md_cities <- md_cities |>
  mutate(
    White_pct = White / TotalPop,
    Black_pct = Black / TotalPop,
    Native_pct = Native / TotalPop,
    Asian_pct = Asian / TotalPop,
    NHPI_pct = NHPI / TotalPop
  )
```

## Task 3: Square the Race Percentages

We need to square the race percentages in order to calculate the probability that any two people chosen at random from a given census area are of different races or are Hispanic or not.

```{r square-race}
md_cities <- md_cities |>
  mutate(
    White_pct_sq = White_pct^2,
    Black_pct_sq = Black_pct^2,
    Native_pct_sq = Native_pct^2,
    Asian_pct_sq = Asian_pct^2,
    NHPI_pct_sq = NHPI_pct^2
  )
```


## Task 4: Calculate Race Probabilities (3 points)

The race probability is the sum the squared probabilities for each of the races. It represents the probability that two people randomly chosen from the same city will be of the same race. Which city has the highest race probability, and why do you think that is?


```{r race-prob}
md_cities <- md_cities |>
  mutate(
    Race_Prob = White_pct_sq + Black_pct_sq + Native_pct_sq + Asian_pct_sq + NHPI_pct_sq
  ) |> 
  arrange(desc(Race_Prob))

View(md_cities)
```

Summerfield has the highest race probability of 0.87. I think the high probability shows that the city has little diversity to it. When looking at the statistics, Summerfield is mostly black people, with very little people from other races leading to the high probability. 

## Task 5: Calculate Ethnicity Percentages and Their Squares (3 points)

Hispanic origin is not a race but an ethnicity or nationality – and a separate Census question. The probability someone is Hispanic or not Hispanic must be calculated separately from race, and Hisp_Prob is the probability that two randomly selected people will be Hispanic or not. That's an important distinction, because unlike race probability, this measures the likelihood of a binary question: that both random people will be Hispanic. What do you notice about the cities with the *lowest* Hisp_Prob scores?


```{r ethnicity}
md_cities <- md_cities |>
  mutate(
    HispTotal = HISPANIC + NotHispanic,
    Hisp_pct = HISPANIC / HispTotal,
    NotHisp_pct = NotHispanic / HispTotal,
    Hisp_pct_sq = Hisp_pct^2,
    NotHisp_pct_sq = NotHisp_pct^2,
    Hisp_Prob = Hisp_pct_sq + NotHisp_pct_sq
  ) |> 
  arrange(Hisp_Prob)

View(md_cities)
```

When looking at Hispanic probability, I noticed that the lower the Hispanic probability, the more diverse the city is. When looking at the hispanic probability values, the cities that correspond to those values have high populations of each race making these ciuties more diverse than others.  

## Task 6: Multiply Race and Ethnicity Probabilities

Here we multiply the two probabilities together to get the probability that any two random people have the SAME race and ethnicity.

```{r combined-prob}
md_cities <- md_cities |>
  mutate(
    RaceEth_Prob = Race_Prob * Hisp_Prob
  )
```


## Task 7: Calculate Diversity Index (3 points)

The final step is to calculate the probability that two people randomly chosen WILL NOT BE OF THE SAME RACE OR ETHNICITY. In other words, the highest index scores will be cities where those two random people will be from different backgrounds. Put simply, more diverse. Which cities are the top 3, and where are they located? Look at the bottom of the list - which city is the least diverse, and where is it?


```{r diversity-index}
md_cities <- md_cities |>
  mutate(
    Diversity_Index = (1 - RaceEth_Prob) * 100
  ) |> 
  arrange(desc(Diversity_Index))

View(md_cities)
```

Top three diverse cities are Langley Park in PG County, Wheaton in Montgomery County and Adelphi in PG County. The least diverse city is Lake Shore in Anne Arundel County

### Pitch a story (8 points)

Given the data that you've produced here, write a headline and lede of the most interesting story from this exercise, and add below that details on how you would report it out: who you'd talk to and the questions you'd ask. When you describe the Diversity Index values, do so in a similar way to how USA Today did; the raw values themselves won't mean much to readers.

Langley Park has the highest Diversity Index in Maryland. 
Recent studies show Langley Park is the most diverse city in Maryland today. 
I would talk to residents in Langley Park and get their opinions on any trends that they notice. Talking to resident will allow me to obtain the best information about any trends of diversity in the city. 
