---
title: "HW7_MandMs"
date: 10/1/2025
---


In this assignment, you'll be picking up where you left off with the M&Ms activity. You'll load the data and perform some basic statistical analysis. Follow the steps below to complete the assignment.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```


### Task 2: Load an M&Ms dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv
```{r}
mandms <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv")
```


### Task 3: Sample 5 rows of data from the dataset, and then sample 10 rows, saving each to separate dataframes called sample5 and sample10. Be sure to use set.seed(1234) before you start.
```{r}
sample_5 <- mandms |> 
  sample_n(5)

sample_10 <- mandms |>
  sample_n(10)

```


### Task 4: Calculate the mean, median and standard deviation of red M&Ms from each of the samples.
```{r}
sample_5 |> summarize(mean(red), median(red), sd(red))
sample_10 |> summarize(mean(red), median(red), sd(red))

```


### Task 5: Calculate the mean, median and standard deviation of red M&Ms from the population

```{r}
mandms |> summarize(mean(red), median(red), sd(red))
```

### Task 6: Make three histograms: one each for sample5, sample10 and the original dataset. Include a headline. Use a consistent binwidth.

```{r}
ggplot(sample_5, aes(x= red, black))
geom_histogram(binwidth = 2, fill = "red", color = "black")
theme_minimal()
labs(title = "5 bag M&M Sample",
     x = "red",
     y = "black")

ggplot(sample_10, aes(x= red, black))
geom_histogram(binwidth = 2, fill = "red", color = "black")
theme_minimal()
labs(title = "10 bag M&M Sample",
     x = "red",
     y = "black")

ggplot(mandms, aes(x= red, black))
geom_histogram(binwidth = 2, fill = "red", color = "black")
theme_minimal()
labs(title = "M&M Distribution",
     x = "red",
     y = "black")
```


### Task 7: Answer questions


Compare the three sets of means, standard deviations and histograms. Which are more helpful to you in understanding the accuracy of the samples?

What does this exercise tell you about the relationship between sample size and estimate accuracy?

How does random sampling introduce variation into statistical estimates?

If you repeated this process, would the samples always be identical? Why or why not?

I think that looking at the mean is the mosat helpful in determining the accuracy of the data sets. It essentially summarizes all the data sets and shows the average number of red M&MS in each bag. This exercise shows that the smaller the sample size is the less accurate the estimation is gonna be because a sample set of 5 or 10 bags is not an accurate representation of the entire population which causes the statistics to be inaccurate. Random sampling includes variation in the data that is given which therefore leads to variations when calculating the statistics fo the data set. Finally, if I were to repeat this process, the samples wouldn't always be identical because it is taking a random sa mple out of the entire population. 

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
