---
title: "Oct 1 Activities: M&Ms; Survey; Sampling"
date: 9-30-2025
---

All three class activities are on this .rmd notebook

Activity #1: Activity: M&Ms - Central Limit Theorem

Working in groups, each group will get a handful of candy from the bag, and then count the number of red and black candies. Log those in this Google Sheet 

https://docs.google.com/spreadsheets/d/1KOeoVTq2W7L2T9sWjwh5U_2f59GsTeAylVGD53rNWww/edit?usp=sharing

(give your group a nice name).

Take a look at the values that we all get - what do you notice about the patterns? This is an exercise about the central limit theorem. 

The central limit theorem states that the sampling distribution of the mean of any independent, random variable will be normal or nearly normal, if the sample size is large enough.

Then discuss and try to answer these questions:

--How close is your group's mean to the true proportion of 0.5?
--Why do different groups get different results?
--What happens to individual extreme results when we combine all groups?
--How does this relate to polls and surveys in real life?

Write your answer below. And then you may eat the M&Ms.

90%
Different groups are getting different results because each group was given different numbers for each color which changed up the results of colored m&ms to brown m&ms. If all groups combined the results the distribution would become more normal because the color of each would be more similar. This relates to polls and surveys in real life because in a survey or poll with multiple answers, there is going to be a different n umber of people that pick each answer, just like the different number of m&ms for each color. 

------------------------------------------------------------------------------

# Activity #2:

Belonging Survey

Working in pairs, take a look back at the main findings from UMD's Belonging Survey:

https://president.umd.edu/internal/belonging


Specifically, focus on the "Sample Description" section and the Appendix. Discuss the differences and between the population and the sample, and what the document says about how the survey attempted to deal with them. Are there any decisions described in the document that raise questions in your minds? Detail the differences you see and your understanding of the methods they employed (it's ok if you don't understand all of them), and add your questions. You should discuss this as a pair but submit individually.

The sample that was used is 23% of the total population. Since there is only 23% of the population that was surveyed, it is not a great representation of the full population that could've been included. The appendix looks normal because it talks about how the survey groups were separated by age, graduate, undergraduate and gender. One question I have is if the dame amount of people from each survey group were surveyed. for example, were the same amount of students and faculty staff included in the survey, and were the same amount of graduate and undergraduate students surveyed. 

------------------------------------------------------------------------------

# TASK 3: SAMPLING

Based on the class sampling demo (available on Elms), sample data from a population using R and the Tidyverse. You should get started by loading the tidyverse in a new R block.

### Task 1: Load the tidyverse

**A**
```{r}
library(tidyverse)

```


### Task 2: Load 1,000 rows of data with the ages of people from 18-80

**A**
```{r}
set.seed(42)
younguns <- tibble(
  age = sample(18:80, 1000, replace = TRUE)
)

head(younguns)

```

### Task 3: Sample 100 rows of data from the dataset
**A**
```{r}
sample_10 <- younguns |> 
  sample_n(10)

sample_50 <- younguns |>
  sample_n(50)

```


### Task 4: Calculate the mean age and standard deviation of the sample
**A**
```{r}
sample_10 |> summarize(mean(age), median(age), sd(age))

```


### Task 5: Calculate the mean age and standard deviation of the population
**A**
```{r}

younguns |> summarize(mean(age), median(age), sd(age))

```


### Task 6: Compare the sample and population statistics


Compare the means and standard deviations. Are they similar? Go back and re-run the sampling and statistics calculations with a smaller sample size. How do the results change?

When comparing the statistics for the sample and the population, the mean and median for the population is higher than the mean and median for the sample. However, the standard deviation of the sample is higher than the standard deviation of the population. when looking at an even smaller sample size, the mean and median increased sign ificantly and ended up being higher than the statistics for the population. 

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
