---
title: "Data Distribution, Volleyball Percentiles, Car Thefts, "
author: "Derek Willis /Rob Wells"
date: 09-23-2025
---

This workbook contains three exercises for our class. Complete them and hand it in at the end of class via a GitHub link.

Your name:
**Answer**
Wesley Schnell and Justin Corman 



# Part 1: When Data Isn't Normal

Working in pairs, consider the following question: "Why do many statistical methods assume a normal distribution, and what challenges arise when data does not meet this assumption?"

First, come up with two examples of data that might not be normally distributed - in other words, that could be skewed or unequal in some way - and explain why they wouldn't be normally distributed. Feel free to draw what you think the distribution would look like.

Then, how should a reporter working on a story about data that isn't normally distributed describe that data? Which statistical measures that we've covered in class should they use or avoid when doing so? Use the examples you came up with, and you should discuss your answers together, but each student submits separately.

**Your answer**

One example of data that might not be normally distributed is a persons income. Incomes can be a wide variety which leads to the data being skewed and not normally distributed. Another example of data that isn't normally distributed is student GPA. If data isn't normally distributed, a reporter can analyze the median because it gives the people the mid point of the data which can help them underatand the data that isn't normally distributed. 

#------------------------------------------------------------------------------#


# Part 2: Volleyball Percentiles

We'll be calculating percentiles to see how good Maryland was in service aces in the 2024 season. We'll use `mutate()` and a new function `ntile()` to calculate percentiles. Get started by loading the tidyverse in a code block, then load the CSV file here: https://raw.githubusercontent.com/dwillis/NCAAWomensVolleyballData/refs/heads/main/data/ncaa_womens_volleyball_matchstats_2024.csv" and save it to a dataframe called `matches`.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```

### Task 2: Load the data

```{r}
matches <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensVolleyballData/refs/heads/main/data/ncaa_womens_volleyball_matchstats_2024.csv")
```

### Task 3: Create totals of aces for each team using `group_by()` and `summarize()`

```{r}
team_aces <- matches |> 
  group_by(team) |> 
  summarize(total_aces = sum(aces)) |> 
  arrange(desc(total_aces))
```


### Task 4: Calculate percentiles using `mutate()` and `ntile()`

```{r}
team_aces |> mutate(percentile = ntile(total_aces, 100))
```


### Task 5: Describe the percentiles

Write a couple of sentences that puts Maryland into context with the other teams in the dataset. You can do some Internet research to help with this.

**Answer**:

For total team aces, Maryland is in the 95th percentile with 223 total aces. This means that Maryland has more aces than 95% of teams that were listed in the table. Towson, for exmaple, is in the 91st percentile for total team aces with 210. Maryland has ore total team aces which shows that they are in the higher percentile.  

#------------------------------------------------------------------------------#

# Part 3: Car Thefts

Follow along with the demo to learn how to calculate standard deviation and how to create a histogram of the data using R and the Tidyverse. Get started by loading the tidyverse in a new R block.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```


### Task 2: Load the car thefts dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/car_thefts_jan24.csv


```{r}
car_thefts <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/car_thefts_jan24.csv")
```

### Task 3: Calculate the standard deviation of the number of car thefts, along with the mean

```{r}
car_thefts |> summarize(mean = mean(count), sd = sd(count))
```


### Task 4: Create a histogram of the number of car thefts


```{r}
car_thefts |> 
  ggplot() + 
  geom_histogram(aes(x = count), binwidth = 2)
```




### Task 5: Add a line to the histogram that shows the mean of the number of car thefts, and two other lines to show one standard deviation in both directions

```{r}
car_thefts |> 
  ggplot() + 
  geom_histogram(aes(x = count), binwidth = 2) +
  geom_vline(aes(xintercept = mean(count)), color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = mean(count) - sd(count)), color = "blue", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = mean(count) + sd(count)), color = "blue", linetype = "dashed", size = 1)
```

### Task 6: Describe your results

Is the distribution normal or skewed in one direction? Write a sentence in which you describe the range of the data covered by one standard deviation from the mean on both sides - how much of the data falls in that range?


**Answer**:
The distribution is skewed to the left. Based on the histogram, about 70% of all the data lies within one standard deviation of the mean. Anything in-between the blue dashed line is one standard deviation away from the mean so about 70% of the data is in-between the blue dashed lines.  





When you are done, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.

Congrats!


